---
title: "Higher Education Salaries Presentation"
subtitle: "Impacts of Education and Demographics on Mid/Early Career Pay"
author: Jonathon Purvis
institute: University Of Scranton
format:
  revealjs:
    self-contained: true
    self_contained-math: true
bibliography: presentation_bib.bib
---

```{r}
#| echo: false
#| include: false

library(tidyverse) # wrangling, plots etc.
library(tidymodels) # modeling
library(skimr) # data summary
library(GGally) # paired plots
library(patchwork) # for combining plots
library(kableExtra) # for extra table formatting 
library(dplyr)

tuition_cost <- read_csv("../data/tuition_cost.csv")
tuition_income <- read_csv("../data/tuition_income.csv")
salary_potential <- read_csv("../data/salary_potential.csv")
historical_tuition <- read_csv("../data/historical_tuition.csv")
diversity_school <- read_csv("../data/diversity_school.csv")
combined_data <- read_csv("../data/combined_data.csv")
states_df <- map_data("state")
```

# Research Question

::: {.incremental}

- Inflation is a popular topic and affects everyone
- Our best weapon against inflation is understanding what factors affect making more money
- **We would like to know impacts of tuition costs, institutional type, STEM percentage, total enrollment, state, and student demographics on mid/early career pay.**

:::

# Tuition Cost Data

```{r}
tuition_cost %>%
  head(2) %>%
  kable()
```

# Tuition Income Data

```{r}
tuition_income %>%
  head(2) %>%
  kable()
```

# Diversity Data

```{r}
diversity_school %>%
  head(2) %>%
  kable()
```

# Salary Data

```{r}
salary_potential %>%
  head(2) %>%
  kable()
```


# Salary Variables

```{r}
#| message: false

early_pay_distribution <- ggplot(salary_potential, aes(x = early_career_pay)) +
  geom_histogram(color = "#512d6d", fill = "lightblue") +
  labs(x = "Early Career Pay (USD)", y = "Count") +
  theme_minimal()

mid_pay_distribution <- ggplot(salary_potential, aes(x = mid_career_pay)) +
  geom_histogram(color = "#512d6d", fill = "lightblue") +
  labs(x = "Mid Career Pay (USD)", y = "Count") +
  theme_minimal()

(mid_pay_distribution + early_pay_distribution)
```

# Tuition Cost Variables

```{r}
#| warning: false

g1 <- ggplot(tuition_cost, aes(x = "", fill = type)) +
  geom_bar(width = 1, color = "white", alpha = 0.7) +
  coord_polar("y") +
  labs(title = "Distribution of Institution Types", fill = "Institution Type") +
  theme_minimal() +
  theme(legend.position = "right")


# Assuming 'data' is your data frame and 'degree_length' is the variable of interest
g2 <- ggplot(tuition_cost, aes(x = "", fill = degree_length)) +
  geom_bar(width = 1, color = "white", alpha = 0.7) +
  coord_polar("y") +
  labs(title = "Distribution of Degree Lengths", fill = "Degree Length") +
  theme_minimal() +
  theme(legend.position = "right")

(g1 + g2)
```

# Tuition Income Variables

```{r}
g3 <- ggplot(tuition_income, aes(x = total_price)) +
  geom_histogram(fill = "lightblue", color = "darkblue", bins = 30, alpha = 0.7) +
  labs(title = "Distribution of Total Price", x = "Total Price", y = "Count") +
  theme_minimal()

g4 <- ggplot(tuition_income, aes(x = net_cost)) +
  geom_histogram(fill = "lightblue", color = "darkblue", bins = 30, alpha = 0.7) +
  labs(title = "Distribution of Net Cost", x = "Net Cost", y = "Count") +
  theme_minimal()

(g3 + g4)
```

# Alumni Percentages Variables

```{r}
#| warning: false

g5 <- salary_potential %>% 
  ggplot(aes(x = stem_percent)) +
  geom_histogram(binwidth = 2, fill = "lightblue", color = "darkblue", alpha = 0.7) +
  labs(title = "Distribution of STEM Percentage", x = "STEM Percentage", y = "Frequency") +
  theme_minimal()

g6 <- salary_potential %>%
  ggplot(aes(x = make_world_better_percent)) +
  geom_histogram(binwidth = 2, fill = "lightblue", color = "darkblue", alpha = 0.7) +
  labs(title = "Distribution of Make World Better Percent", 
       x = "'Make World Better' Percentage", y = "Frequency") +
  theme_minimal()

(g5 + g6)
```

# Total Enrollment

```{r}
filtered_data <- diversity_school %>%
  filter(total_enrollment > 0)

filtered_data

# Histogram for total enrollment
ggplot(filtered_data, aes(x = log(total_enrollment))) +
  geom_histogram(bins = 30, fill = "lightblue", color = "darkblue",boundary = 1) +
  labs(title = "Histogram of Total Enrollment", x = "Total Enrollment", y = "Frequency") +
  theme_minimal()
```

# Diversity

```{r}
ggplot(diversity_school, aes(x = category, y = enrollment)) +
  geom_bar(stat = "identity", color = "lightblue", alpha = 0.7) +
  labs(title = "Enrollment by Category", x = "Category", y = "Enrollment Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


# Variable Relationships

```{r}
# Make world better percent, mid pay
mbp <- ggplot(salary_potential, aes(x = make_world_better_percent, y = mid_career_pay)) +
  geom_point(data = subset(salary_potential, !is.na(make_world_better_percent))) +
  geom_smooth(color = "orange") + 
  labs(
       x = "% of Alumni Making the World Better",
       y = "Mid Career Pay (USD)")


# Stem percent, mid pay
spp <- ggplot(salary_potential, aes(x = stem_percent, y = mid_career_pay)) +
  geom_point() +
  geom_smooth(color = "orange") +  
  labs(x = "STEM Percentage", y = "Mid Career Pay (USD)")

(mbp + spp)
```

# Relationships (Continued)

```{r}
# Merge salary potential and tuition cost
salary_cost <- merge(salary_potential, tuition_cost, by = "name", all.x = TRUE)

# Type, early pay
t1 <- ggplot(filter(salary_cost, !is.na(type) & early_career_pay), 
       aes(x = type, y = early_career_pay)) +
      geom_boxplot() +
      labs(x = "Institution Type", y = "Early Career Pay (USD)")


# Type, mid pay
t2 <- ggplot(filter(salary_cost, !is.na(type) & mid_career_pay), 
       aes(x = type, y = mid_career_pay)) +
  geom_boxplot() +
  labs(x = "Institution Type", y = "Mid Career Pay (USD)")


# Degree length, early pay
dl1 <- ggplot(filter(salary_cost, !is.na(degree_length) & early_career_pay), 
       aes(x = degree_length, y = early_career_pay)) +
  geom_boxplot() +
  labs(x = "Degree Length", y = "Early Career Pay (USD)")


# Degree length, mid pay
dl2 <- ggplot(filter(salary_cost, !is.na(degree_length) & mid_career_pay), 
       aes(x = degree_length, y = mid_career_pay)) +
  geom_boxplot() +
  labs(x = "Degree Length", y = "Mid Career Pay (USD)")

(t1 + t2) / (dl1 + dl2)
```

# Relationships Final

```{r}
# Standardize state names
states_df <- states_df %>%
  mutate(region = tolower(gsub("[_ ]", "", region)))

salary_potential_cleaned <- salary_potential %>%
  mutate(state_name = tolower(gsub("[_ ]", "", state_name))) %>%
  mutate(state_name = gsub("-", "", state_name))

# Left join the data sets, keeping all state boundaries
merged_data <- left_join(states_df, salary_potential_cleaned, by = c("region" = "state_name"))

mid_pay_chloropleth <- merged_data %>%
  ggplot() +
  geom_polygon(aes(long, lat, group = group, fill = mid_career_pay), color = "darkgray") +
  scale_fill_gradient(low = "lightblue", high = "darkblue", labels = scales::comma) +  
  labs(title = "Mid Career Pay by State", fill = "Mid Career Pay") +
  theme_minimal()

mid_pay_chloropleth
```

# Models

```{r}
combined_data <- combined_data %>%
  mutate(log_early_career_pay = log(early_career_pay),
         log_mid_career_pay = log(mid_career_pay))

# Model for pay ~ 
combined_data_split <- initial_split(combined_data)
combined_data_train <- training(combined_data_split)
combined_data_test <- testing(combined_data_split)

# Mid Pay by everything
# ---------
mid_pay_intervals <- reg_intervals(log_mid_career_pay ~ state + total_enrollment + enrollment + 
      total_price + year + net_cost + make_world_better_percent + stem_percent, data= combined_data_train,
                                  keep_reps = TRUE)

p_mid_coeffs <- mid_pay_intervals %>%
  ggplot(aes(x=.estimate,y=term)) + 
  geom_point(size=2) + 
  geom_errorbar(aes(xmin=.lower,xmax=.upper),width=0.1) + 
  geom_vline(xintercept = 0.0,linetype = "dashed") +
  labs(x="Coefficient Estimate",y="Predictor")

p_mid_coeffs
```

# Model (Continued)

```{r}
mid_pay_intervals_2 <- reg_intervals(log_mid_career_pay ~ total_enrollment + enrollment + 
      total_price + year + net_cost + make_world_better_percent + stem_percent, data= combined_data_train,
                                  keep_reps = TRUE)

p_mid_coeffs_2 <- mid_pay_intervals_2 %>%
  ggplot(aes(x=.estimate,y=term)) + 
  geom_point(size=2) + 
  geom_errorbar(aes(xmin=.lower,xmax=.upper),width=0.1) + 
  geom_vline(xintercept = 0.0,linetype = "dashed") +
  labs(x="Coefficient Estimate",y="Predictor")

p_mid_coeffs_2
```


# Conclusions

- Evidence that state, stem percent, and percentage of alumni who think they are making the world a better place influence career salary

- More state data would help with more accurate results









